---
title: "BE_PCLT_report"
author: "Lei Shi"
date: '2022-08-17'
output: 
  pdf_document:
    toc: true
---

```{r message=FALSE, include=FALSE}
# Input everything
library("dplyr")
options(dplyr.summarise.inform = FALSE)

library("ggplot2")
library("tidyverse")
if(!require("AlgDesign")){
  install.packages("AlgDesign")
}
library("AlgDesign")
if(!require("coop")){
  install.packages("coop")
}
library("coop")
library("car")
library("glmnet")

if (!require('latex2exp', quietly = TRUE)) { install.packages('latex2exp') } 
library('latex2exp') # Load package 'Rcpp'

if (!require('Rcpp', quietly = TRUE)) { install.packages('Rcpp') } 
library('Rcpp') # Load package 'Rcpp'
 
if (!require('RcppArmadillo', quietly = TRUE)) { install.packages('RcppArmadillo') } 
library('RcppArmadillo') # Load package 'RcppArmadillo'

library("knitr")

if (!require('qwraps2', quietly = TRUE)) { install.packages('qwraps2') } 
library('qwraps2') # Load package 'qwraps2'

if (!require('xtable', quietly = TRUE)) { install.packages('xtable') } 
library('xtable') # Load package 'qwraps2'

if (!require('matrixStats', quietly = TRUE)) { install.packages('matrixStats') } 
library('matrixStats') # Load package 'qwraps2'

setwd("~/Desktop/Research/BE_PCLT")
source("auxillary_functions.R")
```



# Small effects

```{r include=FALSE}
# generate a factorial experiment
# basic parameter setup
K <- 10
num_trt_group <- 2^K

# design.core <- factor.design(K, trt_group_size = rep(1,2^K), interaction = 2, centering = 1/2)
data.ind <- factor.design(K, trt_group_size = rep(1,2^K), 1)
# design.run <- factor.design(K, trt_group_size = trt_group_size, interaction = 2, centering = 1/2)

active_facs <- rowSums(data.ind)
trt_group_size <- rep(1, 2^K)
trt_group_size[1:660] <- 1
trt_group_size[661:1010]  <- 2
trt_group_size[1011:1024] <- 30
num_pop <- sum(trt_group_size)

Q_U <- 660 # unreplicated small arms
Q_R <- 350  # replicated small arms
Q_L <- 14 # large arms
```



```{r include=FALSE}
# generate factorial design
design.core <- factor.design(K, trt_group_size = rep(1,2^K), interaction = 2, centering = 1/2)
data.ind <- factor.design(K, trt_group_size = trt_group_size, 1)
design.run <- factor.design(K, trt_group_size = trt_group_size, interaction = 2, centering = 1/2)


# generate factorial effects
set.seed(2022)

tau <- c(
  runif(choose(K,0), 0.1, 0.5) * sign(runif(choose(K,0), -1, 1)),
  runif(choose(K,1), 0.05, 0.25) * sign(runif(choose(K,1), -1, 1)),
  runif(choose(K,2), 0.05, 0.25) * sign(runif(choose(K,2), -1, 1)) * rbinom(choose(K,2), 1, 0.7)
)

names(tau) <- c("(Intecept)", names(design.core))

# non-zeros
nonzero.effect <- paste0("F", setdiff(1:K,c(1,4,7,10)))
# nonzero.effect <- paste0("F", 1:2)
nonzero.effect <- c(nonzero.effect, 
                    heredity.proceed.new(K, nonzero.effect, 1,
                                         "strong")$working_model)
zero.effect <- setdiff(names(tau), nonzero.effect)
tau[zero.effect] <- 0

tau
```


```{r include=FALSE}
# generate finite population
mu <- as.matrix(design.core) %*% tau[-1]

finite.pop.opts <- list()

# finite.pop initialization
finite.pop.opts$dist <- rep("exp", 2^K)
finite.pop.opts$mu  <- mu
finite.pop.opts$rate  <- rbinom(2^K, 1, 0.5) + 1
finite.pop.init <- list(
  num_pop = num_pop,
  num_trt_group = num_trt_group,
  finite.pop.opts = finite.pop.opts
)

# generate a finite population
pop_1 <- finite.pop(num_pop, num_trt_group, finite.pop.opts)

pop_cov <- coop::covar(pop_1) 
# very fast!! so much faster than cov()

true_cov_Yhat <- diag(diag(pop_cov)/trt_group_size) - (1/num_pop) * pop_cov

```


```{r include=FALSE}
# specify target effects and target design
target_effect  <- c('F2', 'F4', 'F6', 'F8', 'F10')
target_design  <- design.core[,target_effect]

target_tau <- 1/2^K * t(as.matrix(target_design)) %*% mu

```


```{r include=FALSE}
# population level:
## tau
cat("tau\n")
target_tau
cat("\n")

## variance for the estimator
cat("true variance\n")
true_cov_tauhat <- t(as.matrix(target_design)) %*% true_cov_Yhat %*% as.matrix(target_design) / 1024^2
true_cov_tauhat
sqrt(diag(true_cov_tauhat))
cat("\n")
```



# Experiment setup

- K = 10, number of factors
- Q_U = 660, number of unreplicated arms, each with $N_q = 1$
- Q_R = 350, number of small replicated arms, each with $N_q = 2$
- Q_L = 14, number of large arms, each with $N_q = 30$.
- N = 1780, population

Generate data such that all the $k$-way ($k\ge 3$) interactions are zero. The effects are small: nonzero effects are randomly generated from $\text{uniform}([-0.5,-0.1]\cup [0.1,0.5])$. 

The nonzero factorial effects in the first two levels:
```{r}
tau[abs(tau)>1e-4]
```



The true target effects and true variance for the WLS estimator:
```{r}
# population level:
## tau
cat("target_tau\n")
target_tau
cat("\n")

## variance for the estimator
cat("true variance\n")
true_cov_tauhat <- t(as.matrix(target_design)) %*% true_cov_Yhat %*% as.matrix(target_design) / 1024^2
true_cov_tauhat
cat("\n")

## True standard deviation
cat("true sd\n")
sqrt(diag(true_cov_tauhat))
cat("\n")
```





Target effects we want to estimate: 'F2', 'F4', 'F6', 'F8', 'F10'

 

# Numeric experiments

We run 1000 MC trials and report:

- histogram of the point estimates
- estimated standard deviation for 3 methods (see later parts on what these methods are)
- 95%-CI coverage

## Distribution of point estimates


```{r}
# report results
record <- readRDS("record_SMALL_EFFECTS_EXPN.RData")
# point estimates
hist_data <- data.frame(
  est = c(t(record$rec_point_est)),
  labs = factor(rep(target_effect, each = 1000), levels = c('F2', 'F4', 'F6', 'F8', 'F10'))
)

summary_data <- data.frame(
  target_tau = target_tau,
  labs = factor(target_effect, levels = c('F2', 'F4', 'F6', 'F8', 'F10'))
)

ggplot(hist_data, aes(x = est)) + 
  facet_wrap(~labs, scales = 'free_x') +
  geom_histogram() + 
  geom_vline(data = summary_data, mapping = aes(xintercept = target_tau), col = 'red') # red lines are the true effects

```


```{r}
# point estimates
hist_data <- data.frame(
  est = c(t(record$rec_point_est)) - rep(target_tau, each = 1000),
  labs = factor(rep(target_effect, each = 1000), levels = c('F2', 'F4', 'F6', 'F8', 'F10'))
)

summary_data <- data.frame(
  target_tau = target_tau,
  labs = factor(target_effect, levels = c('F2', 'F4', 'F6', 'F8', 'F10'))
)

plot_SE <- ggplot(hist_data, aes(x = labs, y = est)) + 
  geom_violin(fill = "grey") + 
  theme(legend.position="none",
        text = element_text(size = 20)) +
  scale_x_discrete(labels = c(
    "F2" = TeX("$F_2$"),
    "F4" = TeX("$F_4$"),
    "F6" = TeX("$F_6$"),
    "F8" = TeX("$F_8$"),
    "F10" = TeX("$F_{10}$")
  )) +
  labs(
    y = expression(hat(gamma) - gamma),
    x = "Target effect"
  )

plot_SE


```

```{r eval=FALSE, include=FALSE}
ggsave(
  "BE_PCLT_SMALL_EFFECTS_EXPN.pdf",
  plot = plot_SE,
  device = "pdf", 
  dpi = 300
)
```



Takeaways:

- CLT holds even the design is highly non-uniform.



 

## Expectation of the sd estimators

We applied three methods for variance estimation:

- ehw_0: wls + hc2 var, with specification: $Y\sim F2+F4+F6+F8+F10$
- ehw_1: wls + hc2 var, with specification; $Y\sim (F2+F4+F6+F8+F10)^2$
- lex:   lexicographical pairing

```{r message=FALSE}
# expectation of sd estimator
print(data.frame(
  true_sd   = sqrt(diag(true_cov_tauhat)),
  ehw_0_sd  = diag(apply(sqrt(record$rec_var_est_ehw_0), MARGIN = c(1,2), sum))/1000,
  ehw_1_sd  = diag(apply(sqrt(record$rec_var_est_ehw_1), MARGIN = c(1,2), sum))/1000,
  lex_sd  = diag(apply(sqrt(record$rec_var_est_lex), MARGIN = c(1,2), sum))/1000
))

```



Takeaways: in the small effect cases,

- wls + ehw: both `ehw_0` and `ehw_1` are robust. Adding two-way interactions gives less conservative variance estimation.

- lex:       `lex` pairing is robust. It works better than `ehw` since there is smaller between group variation. 

 

## CI coverage

```{r}
# CI coverage
summary_data <- data.frame(
  lex_coverage  = rowSums(record$rec_coverage_lex)/1000,
  ehw_0_coverage  = rowSums(record$rec_coverage_ehw_0)/1000,
  ehw_1_coverage  = rowSums(record$rec_coverage_ehw_1)/1000
)



summary_data

```





## Power

```{r}
print(data.frame(
  # target_effect = target_effect,
  lex_coverage  = rowSums(record$rec_reject_lex)/1000,
  ehw_0_coverage  = rowSums(record$rec_reject_ehw_0)/1000,
  ehw_1_coverage  = rowSums(record$rec_reject_ehw_1)/1000
))
```






# Large effects



```{r include=FALSE}
# generate a factorial experiment
# basic parameter setup
K <- 10
num_trt_group <- 2^K

# design.core <- factor.design(K, trt_group_size = rep(1,2^K), interaction = 2, centering = 1/2)
data.ind <- factor.design(K, trt_group_size = rep(1,2^K), 1)
# design.run <- factor.design(K, trt_group_size = trt_group_size, interaction = 2, centering = 1/2)

active_facs <- rowSums(data.ind)
trt_group_size <- rep(1, 2^K)
trt_group_size[1:660] <- 1
trt_group_size[661:1010]  <- 2
trt_group_size[1011:1024] <- 30
num_pop <- sum(trt_group_size)

Q_U <- 660 # unreplicated small arms
Q_R <- 350  # replicated small arms
Q_L <- 14 # large arms
```



```{r include=FALSE}
# generate factorial design
design.core <- factor.design(K, trt_group_size = rep(1,2^K), interaction = 2, centering = 1/2)
data.ind <- factor.design(K, trt_group_size = trt_group_size, 1)
design.run <- factor.design(K, trt_group_size = trt_group_size, interaction = 2, centering = 1/2)


# generate factorial effects
set.seed(2023)

tau <- c(
  runif(choose(K,0), 0.1, 0.5) * sign(runif(choose(K,0), -1, 1)),
  runif(choose(K,1), 0.1, 0.5) * sign(runif(choose(K,1), -1, 1)),
  runif(choose(K,2), 0.05, 0.25) * sign(runif(choose(K,2), -1, 1)) * rbinom(choose(K,2), 1, 0.7)
)

names(tau) <- c("(Intecept)", names(design.core))

# non-zeros
nonzero.effect <- paste0("F", setdiff(1:K,c(1,4,7,10)))
# nonzero.effect <- paste0("F", 1:2)
nonzero.effect <- c(nonzero.effect, 
                    heredity.proceed.new(K, nonzero.effect, 1,
                                         "strong")$working_model)
zero.effect <- setdiff(names(tau), nonzero.effect)
tau[zero.effect] <- 0

tau
```


```{r include=FALSE}
# generate finite population
mu <- as.matrix(design.core) %*% tau[-1]

finite.pop.opts <- list()

# finite.pop initialization
finite.pop.opts$dist  <- rep("exp", 2^K)
finite.pop.opts$mu    <- mu
finite.pop.opts$rate  <- rbinom(2^K, 1, 0.5) + 1
finite.pop.init <- list(
  num_pop = num_pop,
  num_trt_group = num_trt_group,
  finite.pop.opts = finite.pop.opts
)

# generate a finite population
pop_1 <- finite.pop(num_pop, num_trt_group, finite.pop.opts)

pop_cov <- coop::covar(pop_1) 
# very fast!! so much faster than cov()

true_cov_Yhat <- diag(diag(pop_cov)/trt_group_size) - (1/num_pop) * pop_cov

```


```{r include=FALSE}
# specify target effects and target design
target_effect  <- c('F2', 'F4', 'F6', 'F8', 'F10')
target_design  <- design.core[,target_effect]

target_tau <- 1/2^K * t(as.matrix(target_design)) %*% mu

```


```{r include=FALSE}
# population level:
## tau
cat("tau\n")
target_tau
cat("\n")

## variance for the estimator
cat("true variance\n")
true_cov_tauhat <- t(as.matrix(target_design)) %*% true_cov_Yhat %*% as.matrix(target_design) / 1024^2
true_cov_tauhat
sqrt(diag(true_cov_tauhat))
cat("\n")
```



# Experiment setup

- K = 10, number of factors
- Q_U = 660, number of unreplicated arms, each with $N_q = 1$
- Q_R = 350, number of small replicated arms, each with $N_q = 2$
- Q_L = 14, number of large arms, each with $N_q = 30$.
- N = 1780, population

Generate data such that all the $k$-way ($k\ge 3$) interactions are zero. The effects are small: nonzero effects are randomly generated from $\text{uniform}([-0.5,-0.1]\cup [0.1,0.5])$. 

The nonzero factorial effects in the first two levels:
```{r}
tau[abs(tau)>1e-4]
```



The true target effects and true variance for the WLS estimator:
```{r}
# population level:
## tau
cat("target_tau\n")
target_tau
cat("\n")

## variance for the estimator
cat("true variance\n")
true_cov_tauhat <- t(as.matrix(target_design)) %*% true_cov_Yhat %*% as.matrix(target_design) / 1024^2
true_cov_tauhat
cat("\n")

## True standard deviation
cat("true sd\n")
sqrt(diag(true_cov_tauhat))
cat("\n")
```





Target effects we want to estimate: 'F2', 'F4', 'F6', 'F8', 'F10'

 

# Numeric experiments

We run 1000 MC trials and report:

- histogram of the point estimates
- estimated standard deviation for 3 methods (see later parts on what these methods are)
- 95%-CI coverage

## Distribution of point estimates


```{r}
# report results
record <- readRDS("record_LARGE_EFFECTS_EXPN.RData")
# point estimates
hist_data <- data.frame(
  est = c(t(record$rec_point_est)),
  labs = factor(rep(target_effect, each = 1000), levels = c('F2', 'F4', 'F6', 'F8', 'F10'))
)

summary_data <- data.frame(
  target_tau = target_tau,
  labs = factor(target_effect, levels = c('F2', 'F4', 'F6', 'F8', 'F10'))
)

ggplot(hist_data, aes(x = est)) + 
  facet_wrap(~labs, scales = 'free_x') +
  geom_histogram() + 
  geom_vline(data = summary_data, mapping = aes(xintercept = target_tau), col = 'red') # red lines are the true effects



```



```{r}
# point estimates
hist_data <- data.frame(
  est = c(t(record$rec_point_est)) - rep(target_tau, each = 1000),
  labs = factor(rep(target_effect, each = 1000), levels = c('F2', 'F4', 'F6', 'F8', 'F10'))
)

summary_data <- data.frame(
  target_tau = target_tau,
  labs = factor(target_effect, levels = c('F2', 'F4', 'F6', 'F8', 'F10'))
)

plot_LE <- ggplot(hist_data, aes(x = labs, y = est)) + 
  geom_violin(fill = "grey") + 
  theme(legend.position="none",
        text = element_text(size = 20)) +
  scale_x_discrete(labels = c(
    "F2" = TeX("$F_2$"),
    "F4" = TeX("$F_4$"),
    "F6" = TeX("$F_6$"),
    "F8" = TeX("$F_8$"),
    "F10" = TeX("$F_{10}$")
  )) +
  labs(
    y = expression(hat(gamma) - gamma),
    x = "Target effect"
  ) 

plot_LE

```


```{r eval=FALSE, include=FALSE}
ggsave(
  "BE_PCLT_LARGE_EFFECTS_EXPN.pdf",
  plot = plot_LE,
  device = "pdf", 
  dpi = 300
)
```


Takeaways:

- CLT holds even the design is highly non-uniform.

 

## Expectation of the sd estimators

We applied three methods for variance estimation:

- ehw_0: wls + hc2 var, with specification: $Y\sim F2+F4+F6+F8+F10$
- ehw_1: wls + hc2 var, with specification; $Y\sim (F2+F4+F6+F8+F10)^2$
- lex:   lexicographical pairing

```{r message=FALSE}
# expectation of sd estimator
print(data.frame(
  true_sd   = sqrt(diag(true_cov_tauhat)),
  ehw_0_sd  = diag(apply(sqrt(record$rec_var_est_ehw_0), MARGIN = c(1,2), sum))/1000,
  ehw_1_sd  = diag(apply(sqrt(record$rec_var_est_ehw_1), MARGIN = c(1,2), sum))/1000,
  lex_sd  = diag(apply(sqrt(record$rec_var_est_lex), MARGIN = c(1,2), sum))/1000
))

```



Takeaways: in the small effect cases,

- wls + ehw: both `ehw_0` and `ehw_1` are robust. Adding two-way interactions gives less conservative variance estimation.

- lex:       `lex` pairing is robust. It works better than `ehw` since there is smaller between group variation. 

 

## CI coverage

```{r}
# CI coverage
data.frame(
  # target_effect = target_effect,
  lex_coverage  = rowSums(record$rec_coverage_lex)/1000,
  ehw_0_coverage  = rowSums(record$rec_coverage_ehw_0)/1000,
  ehw_1_coverage  = rowSums(record$rec_coverage_ehw_1)/1000
)

```


```{r}

```



## Power

```{r}
print(data.frame(
  #target_effect = target_effect,
  lex_coverage  = rowSums(record$rec_reject_lex)/1000,
  ehw_0_coverage  = rowSums(record$rec_reject_ehw_0)/1000,
  ehw_1_coverage  = rowSums(record$rec_reject_ehw_1)/1000
))
```







