---
title: "BE_PCLT_report"
author: "Lei Shi"
date: '2022-08-17'
output: pdf_document
---

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
```


## Experiment setup

- K = 10, number of factors
- Q_U = 638, number of unreplicated arms, each with $N_q = 1$
- Q_R = 375, number of small replicated arms, each with $N_q = 2$
- Q_L = 11, number of large arms, each with $N_q = 30$.
- N = 1718, population

Generate data such that all the $k$-way ($k\ge 3$) interactions are zero.

Target effects we want to estimate: 'F1', 'F2', 'F3', 'F1.F2', 'F1.F3', 'F2.F3'


## Experiment with correct specifications

In this experiments, we run 1000 MC trials. The data are generated such that only 'F1', 'F2' and 'F1.F2' have nonzero effects.


```{r}
# report results
record <- readRDS("record_CORRECT_SPEC.RData")
target_effect  <- c('F1', 'F2', 'F3', 'F1.F2', 'F1.F3', 'F2.F3')
# point estimates
hist_data <- data.frame(
  est = c(t(record[[1]])),
  labs = factor(rep(target_effect, each = 1000), levels = c('F1', 'F2', 'F3', 'F1.F2', 'F1.F3', 'F2.F3'))
)

ggplot(hist_data, aes(x = est)) + 
  facet_wrap(~labs, scales = 'free_x') +
  geom_histogram()

# true effects: -1.481314  3.175201  0.000000 -2.706229  0.000000  0.000000
```

```{r}
# expectation of sd estimator
target_effect  <- c('F1', 'F2', 'F3', 'F1.F2', 'F1.F3', 'F2.F3')
print(data.frame(
  target_effect = target_effect,
  wls_coverage  = diag(apply(sqrt(record[[2]]), MARGIN = c(1,2), sum))/1000,
  ehw_coverage  = diag(apply(sqrt(record[[3]]), MARGIN = c(1,2), sum))/1000,
  lex_coverage  = diag(apply(sqrt(record[[4]]), MARGIN = c(1,2), sum))/1000
))

# true sd's:
# 0.04368610 0.04368584 0.04368498 0.04368678 0.04368549 0.04368564
```



```{r}
# CI coverage
target_effect  <- c('F1', 'F2', 'F3', 'F1.F2', 'F1.F3', 'F2.F3')
print(data.frame(
  target_effect = target_effect,
  wls_coverage  = rowSums(record[[5]])/1000,
  ehw_coverage  = rowSums(record[[6]])/1000,
  lex_coverage  = rowSums(record[[7]])/1000
))

```


```{r}
# wald inference
print(data.frame(
  method = c('WLS', 'EHW', 'LEX'),
  wald_coverage = c(sum(record[[8]])/1000, sum(record[[9]])/1000, sum(record[[10]])/1000)
))
```




## Experiment with incorrect specifications

In this experiments, we run 1000 MC trials. The data are generated such that only 'F1', 'F3', 'F5', 'F7', 'F9' have nonzero main effects. The two-way interactions are generated through strong heredity so some of them are also nonzero.

```{r}
# report results
record <- readRDS("record_INCORRECT_SPEC.RData")
target_effect  <- c('F1', 'F2', 'F3', 'F1.F2', 'F1.F3', 'F2.F3')
# point estimates
hist_data <- data.frame(
  est = c(t(record[[1]])),
  labs = factor(rep(target_effect, each = 1000), levels = c('F1', 'F2', 'F3', 'F1.F2', 'F1.F3', 'F2.F3'))
)

ggplot(hist_data, aes(x = est)) + 
  facet_wrap(~labs, scales = 'free_x') +
  geom_histogram()

# true effects: -1.481314  0.000000  1.738920  0.000000 -1.548822  0.000000
```

```{r}
# expectation of sd estimator
target_effect  <- c('F1', 'F2', 'F3', 'F1.F2', 'F1.F3', 'F2.F3')
print(data.frame(
  target_effect = target_effect,
  wls_coverage  = diag(apply(sqrt(record[[2]]), MARGIN = c(1,2), sum))/1000,
  ehw_coverage  = diag(apply(sqrt(record[[3]]), MARGIN = c(1,2), sum))/1000,
  lex_coverage  = diag(apply(sqrt(record[[4]]), MARGIN = c(1,2), sum))/1000
))

# true sd's:
# 0.04368610 0.04368584 0.04368498 0.04368678 0.04368549 0.04368564 
```



```{r}
# CI coverage
target_effect  <- c('F1', 'F2', 'F3', 'F1.F2', 'F1.F3', 'F2.F3')
print(data.frame(
  target_effect = target_effect,
  wls_coverage  = rowSums(record[[5]])/1000,
  ehw_coverage  = rowSums(record[[6]])/1000,
  lex_coverage  = rowSums(record[[7]])/1000
))

```


```{r}
# wald inference
print(data.frame(
  method = c('WLS', 'EHW', 'LEX'),
  wald_coverage = c(sum(record[[8]])/1000, sum(record[[9]])/1000, sum(record[[10]])/1000)
))
```